Beyesian Inference for Decision Making under COVID-19: A case study

### 2020-05-29 out/paper.Rmd
Draft sección 2

Although the normal parametrization is convenient for its simplicity, it should stand out to any statistician that there is a major problem choosing the gaussian family because $\theta$ $\epsilon$ $\Theta = [ 0, \infty ]$ and under the normality assumption $\Theta =  [- \infty, \infty]$. For now, our way around will be to truncate such $N(\hat{\theta}, \hat{\sigma^2})$ and redistribute tail density [^1].

@@

Many coleagues would argue that this textbook like toy model doesn´t even correspond to ordered nature of observations or even the parametral space. However any other, more sophisticated, model would encounter the same incovenience as there would be no input accounting for COVID-19 induced contraction at the time.

@@

  b. Katia, a fellow friend of Marcos, suggests increasing $\sigma_{MLE}^2$ by some amount to reflect the uncertainty. Marcos agrees with katia, however the question that will innevitably arise is, by how much?.

@@

#

>>>>

Although the normal parametrization is convenient for its simplicity, it should stand out to anyone that there is a major problem choosing the gaussian family because $\theta$ $\epsilon$ $\Theta = [ 0, \infty ]$ and under the normality assumption $\Theta =  [- \infty, \infty]$. For now, our way around will be to truncate such $N(\hat{\theta}, \hat{\sigma^2})$ and redistribute tail density [^1].

@@

Many coleagues would argue that this textbook like toy model doesn´t even correspond to ordered nature of observations or even the parametral space. However any other, more sophisticated, model would encounter the same incovenience as there would be no input accounting for COVID-19 induced contraction at the time, not to mention the rise in model complexity and the need for computer power.

@@

  b. Katia, a fellow colleague of Marcos, suggests increasing $\sigma_{MLE}^2$ by some amount to reflect the uncertainty. Marcos agrees with katia, however the question that will innevitably arise is, by how much?.

@@

# Incorporating reverend Bayes square table

Rescuing our "not-so-toy" example, an alternative to adress some of the classical framework caveats is to proceed through bayesian grounds. We'll part from the same assumptions and the same structural construction from the model except, for noew, we will asume variance to be known and equal to historical variance, additionally we assume a reference distribution for $\theta$ given by $N(\theta | \mu_0, \sigma_0^2)$ in orider to take advantage of a conjugacy scheme. Building the model this way, we get that the posterior predictive distribution for $\theta_{t+1} = \theta_{COVID-19}$ is given by $$f(X_{t+1}| X_{(\underline{n})}) \sim N(x | \mu_{n}, \sigma_{n}^2 + \sigma_{x}^2)$$

Where $\mu_n = \sigma_n^2 \cdot (\frac{\mu_0}{\sigma_0^2}+\frac{n \overline{x}}{\sigma_x^2})$ and $\sigma_n^2 = \frac{1}{\frac{n}{\sigma_x^2}+ \frac{1}{\sigma_0^2}}$.

Commenting on the forgoing model, it could de said that we gain complexity at the cost os assuming a known variance, which in obviously not true. Nevertheless we gained so much more more than a little mathematical complexity, this being a consistent way to introduce Marco's expert knowledge into the inferential process as an endogenous input through $\theta$ 's *a priori* distribution.

Following this line of though, Marco's expert knowledge is introduced in the form of a (highly) **informative prior**. Expertise is incorporated first by centering the future bahaviour of the lagged number of ployees around 5% and second by adjusting the level of "surenees" by declaring a very low standard deviation [^3] lets say .001%. In fact chossing this particular characterization for $\theta_{COVID-19}$ 's prior yields an prediction credibility interval of **INSERTAR INTERVALO DE CREDIBILIDAD** around **INSERTAR ESTIMACIÓN PUNTUAL** [^4].

[^3]: Conversely with very high precision $\tau = \frac{1}{\sigma^2}$

[^4]: After truncation

Even though this specific prior characterization is as arbitrary as any number choosen for a *mark-down procedure* the *a prior* and not *a posteriori* fashion of introducing this knowledge, plus the interaction between paramters through the conjugacy scheme and the **likelihood resilience** [^5] allow for a coherent manner to let hypothesis and data interact with each other parsimonious fashion within the walls of a model[^6].

[^5]: Haro-Peniche 2020
[^6]: This also adresses concern **b** from section I

From last paragraph its particuarly worth discussing the term *likelihood resilience* from the known variance Normal-Normal conjugate family. Which is formally measured as the relative in change in the posterior distributions parameters in the presence of extreme prioir **FALTA PULIR ESA DEFINICIÓN CON MEDIDAS RELATIVAS ESPECÍFICAS EN TERMINOS DE MOVIMIENTOS EXTREMOS DE LA PRIOR PARA CIERTO TRHESHOLD**. In Figure 1 we can aprreciate the posteriors distribution behaviour with different prior paramters.

**FIGURE ONE**

**DESCRIBIR COMPORTAMIENTO**. In other words likelihood resilience in this model is not just an intrinsic characteristic that prevents arbitrary data coercion or the *lying with statistics* issue but a quantifiable measure of how robust the inference is against extreme hypotheis.

Nonetheless many may be still be troubled by the known variance assumption, which is in fact totally reasonable. Relaxing this premise is just a matter of a litter more mathematical intricacy [^7]. For a more realistic scenario we choose the Normal-InverseGamma family where the predictive distribution for $\hat{\theta}_{COVID-19}$ is: $$f(X_{na+1} | X_{(\underline{n})}) \sim  t_{2 \alpha_n}(x | \mu_n, \frac{\beta_n( \kappa_n + 1)}{\alpha_n \kappa_n})$$
Where:

  + $\mu_n = \frac{\kappa_0  \mu_0 + n \overline{x}}{\kappa_0 + n}$
  + $\beta_n = \beta_0 + \frac{1}{2} \sum_{i=1}^n (x_i - \overline{x})^2 + \frac{\kappa_0 n (\overline{x} - \mu_0)}{2(\kappa_0 + n)}$
  + $\kappa_n = \kappa_0 + n$
  + $\alpha_n = \alpha_0 + \frac{n}{2}$

Additionally the posterior ma

[^7]: All normal conjugacy cases are weel documented in **Insertar cita**
